### 2021-09-03 LSOA domestic Elec and Gas.Rmd
restructured headings and order so the headings work in html; load boundaries & deprivation data at the start; re-use loaded data where we can; re-order of plots slightly; reduce all dataframes so we just merge in the variables we want; more informative captions

---
title: "LSOA domestic Electricty and Gas maps"
author: "Meghan Kinglsey-Walsh"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # set this to FALSE if you don't want the code in the final output

# useful R libraries ----
library(dplyr)
library(ggplot2)
library(leaflet)
library(readr)

# GIS libraries ----
library(raster)
library(sf)
#library(spData) # not needed
#library(spDataLarge)

# local functions ----
source(here::here("R", "functions.R"))

dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/" # default for Meghan
user <- Sys.info()[[7]] # who are we logged in as?
if(user == "ben"){
  dataFolder <- path.expand("~/University of Southampton/HCC Energy Landscape Mapping project - Documents/General/data/")
}

# this will not print anything out when knitted due to the include=FALSE setting in the chunk header
message("User: ", user)
message("dataFolder: ", dataFolder)
```

# Background data

Load & check boundary data for Solent area. The table shows which LAs we expect to plot.

```{r Loading in boundary data for the LSOA}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "boundaries/LSOA/lsoa_solent.shp")
message ("Loading LSOA boundaries from file")
lsoa_sf <- sf::read_sf(inf)
head(lsoa_sf)
table(lsoa_sf$LAD11NM) # How many LSOAs are there in each LA?
```

```{r transform boundary data for use in leaflet}
st_coord_sys <- st_crs(lsoa_sf)
st_coord_sys$epsg

#Trasforming the coord systen
if(st_coord_sys$epsg !=4326){
  lsoa_sf_leaflet <- st_transform(lsoa_sf, "+proj=longlat +datum=WGS84")
}
```

Loading deprivation data and double checking what the IMD Decile means - NB this is all LSOAs not just Solent

```{r Loading in the Deprivation score data}
inFile<-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv")
LSOAdep<-readr::read_csv(inFile)
LSOAdep$'LSOA11CD' <- LSOAdep$'lsoa11cd'
head(LSOAdep)
#names(LSOAdep)

ggplot(LSOAdep, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y = IMDScore, 
                   group = IMD_Decile,
                   colour = IMD_Decile)) +  
  geom_boxplot() +
  scale_color_continuous(name = "IMD Decile") + # rename the legend scale
  labs(x = "IMD Decile",
       y = "IMD Score",
       caption = "All LSOAs")

LSOAdep$la_name <- LSOAdep$LADnm
solent_dep <- getSolent(LSOAdep)
ggplot(solent_dep, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y = IMDScore, 
                   group = IMD_Decile,
                   colour = IMD_Decile)) +  
  geom_boxplot() +
  scale_color_continuous(name = "IMD Decile") + # rename the legend scale
  facet_wrap(. ~ la_name)
  labs(x = "IMD Decile",
       y = "IMD Score",
       caption = "Solent LSOAs split by Local Authority")
```

> NB: IMD Decile 10 has _LOWER_ mean IMD score (less deprived)! Which is what it says in the Excel workbook...

> So remember: IMD Decile 1 = 10% _most_ deprived; IMD Decile 10 = 10% _least_ deprived

# Domestic Electricity consumption

## 2019

Load LSOA elec data - NB this is all LSOAs not just Solent

```{r Loading in the LSOA electricty consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/electricity/LSOA Dom Elec csv/LSOA_ELEC_2019.csv")
domelec_LSOA_2019 <- readr::read_csv(inFile)
domelec_LSOA_2019$LSOA11CD <- domelec_LSOA_2019$`Lower Layer Super Output Area (LSOA) Code`
head(domelec_LSOA_2019)
domelec_LSOA_2019$mean_de2019 <- domelec_LSOA_2019$`Mean domestic electricity consumption 
(kWh per meter)`
```

### Base map

```{r Mapping the electricty consumption of the LSOAs}
# just add the variables we want
df <- dplyr::select(domelec_LSOA_2019, LSOA11CD, mean_de2019)

lsoa_merged_sf <- merge(lsoa_sf, df) #merging the LSOA boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf) + 
  geom_sf(aes(fill = mean_de2019)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Electricity (2019, all Solent LSOAs)")
#Mapping it
```

### Map with Overlay (leaflet)

```{r Plotting leaflet dom electricity consumption map}
# make a new df for use in the map (prevents confusion)
map_df <- merge(lsoa_sf_leaflet, df, by = "LSOA11CD")

qpal <- colorNumeric("Reds", df$mean_de2019, n=9)

leaflet(map_df) %>%
addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(color = ~qpal(mean_de2019),
              fillOpacity = 0.6, weight = 1.5, popup = ~(LSOA11CD),

               highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```
### Looking at cities in more detail

```{r Southampton LSOA dom Elec}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf, LAD11NM == "Southampton")
# plotting the electricity consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_de2019)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Southampton")
```

```{r Portsmouth LSOA dom Elec}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf, LAD11NM == "Portsmouth")
# plotting the electricity consumption of Portsmouth in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_de2019)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Portsmouth")
```

```{r Winchester LSOA dom Elec}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf, LAD11NM == "Winchester")

# plotting the electricity consumption of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_de2019)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Winchester")
```

### Plotting the relationship between Electricty consumption and deprivation

To see if higher electricity use is associated with higher deprivation

```{r plotting electricty consumption against deprevation}
domelec_LSOA_2019$la_name <-domelec_LSOA_2019$`Local Authority Name`

solent_elec <- getSolent(domelec_LSOA_2019)

df <- merge(solent_dep, solent_elec, by = "LSOA11CD")

ggplot(df, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y = mean_de2019, 
                   group = IMD_Decile,
                   colour = IMD_Decile)) +  
  geom_boxplot() +
  scale_color_continuous(name = "IMD Decile") + # rename the legend scale
  labs(x = "IMD Decile",
       y = "Electricity use (mean kWh per meter)",
       caption = "Solent LSOAs, 2019")
```

With facets by LA

```{r addFacetsToElecconsumption}
# or without colour but using facets for Local Authorities
ggplot(df, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y =mean_de2019, 
                   group = IMD_Decile)) +
  geom_boxplot() +
  labs(x = "IMD Decile",
       y = "Electricity use (mean kWh per meter)",
       caption = "Solent LSOAs, 2019") +
  facet_wrap(la_name.x ~ .)

```

### Difference between 2010 and 2019 dom elec

```{r Creating the difference in LSOA mean dom elec consumption variable}
domelec_LSOA_2010 <- readr::read_csv(paste0(dataFolder,"energy/electricity/LSOA Dom Elec csv/LSOA_ELEC_2010.csv"))
domelec_LSOA_2010$mean_de2010 <- domelec_LSOA_2010$'Mean domestic electricity consumption \n(kWh per meter)'

# create unambiguous lsoa code for matching to each other and the sf geometry
domelec_LSOA_2010$LSOA11CD <- domelec_LSOA_2010$`Lower Layer Super Output Area (LSOA) Code`
# just the vars we want
df <- dplyr::select(domelec_LSOA_2010, LSOA11CD, mean_de2010)
# reuse the 2019 data we already analysed
merged_de <- merge(df, domelec_LSOA_2019, by="LSOA11CD", all=TRUE)
#names(merged_de)

merged_de$mean_de_diff <- as.numeric(merged_de$mean_de2019)-as.numeric(merged_de$mean_de2010) 
merged_de$mean_de_diff_pc <- 100*(merged_de$mean_de_diff/merged_de$mean_de2010)
summary(merged_de$mean_de_diff)
summary(merged_de$mean_de_diff_pc)
```

#### Basic map

```{r Mapping the LSOA difference in mean dom elec consumption}
# merge to sf geometry on LSOA11CD
lsoa_sf_de20102019<-merge(lsoa_sf, merged_de, by = "LSOA11CD")
ggplot2::ggplot(lsoa_sf_de20102019) +
  geom_sf(aes(fill=mean_de_diff))+
  scale_fill_continuous(name="Change in Mean consumpton (kWh per meter)",low="green",high="red") +
  labs(caption="Solent (all LSOAs)")

t_diff <-  dplyr::select(as.data.frame(lsoa_sf_de20102019), 
                         LSOA11CD,
                         mean_de2010, mean_de2019, mean_de_diff, mean_de_diff_pc)
head(arrange(t_diff, mean_de_diff_pc))
head(arrange(t_diff, -mean_de_diff_pc))
ggplot2::ggplot(t_diff, aes(x = mean_de_diff_pc)) +
  geom_histogram()
```

#### Change map with Overlay (leaflet)

```{r Plotting leaflet for change in dom elec consumption map}
# merge just the variables we want to the leaflet sf
df <- dplyr::select(merged_de, LSOA11CD, mean_de2010, mean_de2019, mean_de_diff, mean_de_diff_pc)
lsoa_sf_leaflet <- merge(lsoa_sf_leaflet, df) # add on to the 
#names(lsoa_sf_leaflet)

qpal <- colorNumeric("Reds", lsoa_sf_leaflet$mean_de_diff, n=9)

# we already checked & transformed the boundaries for leaflet so re-use
leaflet(lsoa_sf_leaflet) %>% 
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(color = ~qpal(mean_de_diff),
              fillOpacity = 0.6, weight = 1.5, popup = ~(LSOA11CD ),

               highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

#### Cities in more detail

```{r Southampton LSOA difference in dom Elec}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_sf_de20102019, LAD11NM == "Southampton")
# plotting the electricity consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill= mean_de_diff)) +
  scale_fill_continuous(name = "Change in Mean consumpton (kWh per meter)", low = "green", high = "red") +
  labs(caption = "Southampton")
#BROKEN
```
```{r Portsmouth LSOA difference in dom Elec}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_sf_de20102019, LAD11NM == "Portsmouth")
# plotting the electricity consumption of Portsmouth in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_de_diff)) +
  scale_fill_continuous(name = "Change in Mean comsumpton (kWh per meter)", low = "green", high = "red") +
  labs(caption = "Portsmouth")
```
```{r Winchester LSOA difference in dom Elec}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_sf_de20102019, LAD11NM == "Winchester")
# plotting the electricity consumption of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_de_diff)) +
  scale_fill_continuous(name = "Change in Mean comsumpton (kWh per meter)", low = "green", high = "red") +
  labs(caption = "Winchester")
```

#### Plotting the change in dom elec consumption against the deprivation score

```{r plotting change in elec usage against dep}
df <- dplyr::select(merged_de, LSOA11CD, mean_de_diff, mean_de2010, mean_de2019, mean_de_diff_pc)
LSOAdepelec <- merge(df, LSOAdep, by= "LSOA11CD")
#names(LSOAdepelec)

Solent <- getSolent(LSOAdepelec)
head(Solent)
ggplot(Solent, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y = mean_de_diff_pc, 
                   group = IMD_Decile,
                   colour = IMD_Decile)) +  
  geom_boxplot() +
  scale_color_continuous(name = "IMD Decile") + # rename the legend scale
  labs(x = "IMD Decile",
       y = "% change in mean kWh per meter",
       caption = "Electricity 2010-2019, all Solent LSOAs")
```

Try adding facets for each Local Authority

```{r addFacetsToElecDiff}
# or without colour but using facets for Local Authorities
ggplot(Solent, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y =mean_de_diff_pc, 
                   group = IMD_Decile)) +
  geom_boxplot() +
  labs(x = "IMD Decile",
       y = "% change in mean kWh per meter",
       caption = "Electricity 2010-2019, all Solent LSOAs") +
  facet_wrap(la_name ~ .)

nrow(Solent)
```

# Mapping Gas

## 2019

```{r Loading in the LSOA gas consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/gas/LSOA Gas csv/LSOA_GAS_2019.csv")
domgas_LSOA_2019 <- readr::read_csv(inFile)
domgas_LSOA_2019$LSOA11CD <- domgas_LSOA_2019$`Lower Layer Super Output Area (LSOA) Code`

domgas_LSOA_2019$mean_dg2019 <- domgas_LSOA_2019$`Mean consumption (kWh per meter)`

head(domgas_LSOA_2019)
```

### Basic map

```{r Mapping the gas consumption of the LSOAs}
df <- dplyr::select(domgas_LSOA_2019, LSOA11CD, mean_dg2019)
lsoa_merged_sf_gas <- merge(lsoa_sf, df) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf_gas) + 
  geom_sf(aes(fill = mean_dg2019)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```
> Aha - this one shows where there is no gas :-)

### Map with Overlay

```{r Plotting leaflet dom gas consumption}
# re-use the df from previous chunk
lsoa_sf_leaflet <- merge(lsoa_sf_leaflet, df, by = "LSOA11CD")
qpal <- colorNumeric("Reds", lsoa_sf_leaflet$mean_dg2019, n=9)

leaflet(lsoa_sf_leaflet) %>%
  addTiles() %>%
  addPolygons(color= ~qpal(mean_dg2019), 
              fillOpacity = 0.6, weight = 1.5, popup = ~(LSOA11CD),

              highlight=highlightOptions(
                weight=5,
                color="#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

### Cities in More detail

```{r Southampton LSOA Gas}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf_gas, LAD11NM == "Southampton")

# plotting the gas consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_dg2019)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Southampton")
```
```{r Portsmouth LSOA Gas}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf_gas, LAD11NM == "Portsmouth")

# plotting the gas consumption of Portsmouth in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_dg2019)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Portsmouth")
```

```{r Winchester LSOA Gas}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf_gas, LAD11NM == "Winchester")

# plotting the gas consumption of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = mean_dg2019)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Winchester")
```

### Plotting  gas usage against deprivation

```{r Boxplot of the gas consumption and deprevation}
domgas_LSOA_2019$la_name <-domgas_LSOA_2019$`Local Authority Name`

solent_gas <- getSolent(domgas_LSOA_2019)

df <- merge(solent_dep, solent_gas, by = "LSOA11CD")

ggplot(df, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y = mean_dg2019, 
                   group = IMD_Decile,
                   colour = IMD_Decile)) +  
  geom_boxplot() +
  scale_color_continuous(name = "IMD Decile") + # rename the legend scale
  labs(x = "IMD Decile",
       y = "Gas Consumption")

```

Much more of a correlation with low deprivation (higher deciles)

```{r addFacetsTogasconsumption}
# or without colour but using facets for Local Authorities
ggplot(df, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y =mean_dg2019, 
                   group = IMD_Decile)) +
  geom_boxplot() +
  labs(x = "IMD Decile",
       y = "Gas Consumption") +
  facet_wrap(la_name.y ~ .)

nrow(Solent)
```

### Difference between 2010 and 2019 dom gas

```{r Creating the dfference in LSOA dom gas consumption variable}

domgas_LSOA_2010 <- readr::read_csv(paste0(dataFolder, "energy/gas/LSOA Gas csv/LSOA_GAS_2010.csv"))
domgas_LSOA_2010$mean_dg2010 <- domgas_LSOA_2010$'Mean consumption (kWh per meter)'
domgas_LSOA_2010$LSOA11CD <- domgas_LSOA_2010$`Lower Layer Super Output Area (LSOA) Code`

df <- dplyr::select(domgas_LSOA_2010, LSOA11CD, mean_dg2010) # just the vars we need

merged_dgas <- merge(domgas_LSOA_2019, df, by = "LSOA11CD", all = TRUE)

#names(merged_dgas)
merged_dgas$dg_diff<-as.numeric(merged_dgas$mean_dg2019) - merged_dgas$mean_dg2010
summary(merged_dgas)
```

#### Plotting the change in gas usage against the deprivation

```{r plotting change in gas usage against dep}
LSOAdepgas <- merge(merged_dgas, LSOAdep, by= "LSOA11CD")
#names(LSOAdepgas)
df <- dplyr::select(LSOAdepgas, LSOA11CD, la_name.x, mean_dg2019, IMDScore, mean_dg2010, IMD_Decile,dg_diff
)

head(df)
df$la_name <- df$la_name.x
Solent <- getSolent(df)
head(Solent)
ggplot(Solent, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y =100*(dg_diff/mean_dg2010), 
                   group = IMD_Decile,
                   colour = IMD_Decile)) +  
  geom_boxplot() +
  scale_color_continuous(name = "IMD Decile") + # rename the legend scale
  labs(x = "IMD Decile",
       y = "% change",
       caption = "Mean gas kWh 2010-2019, all Solent LSOas")
```

Try adding facets for each Local Authority and remove the outlier

```{r addFacetsTogasDiff}
# or without colour but using facets for Local Authorities
Solentex <- filter(Solent,100*(dg_diff/mean_dg2010)<50)
ggplot(Solentex, aes(x = as.factor(IMD_Decile), # forces a box plot at each value 
                   y =100*(dg_diff/mean_dg2010), 
                   group = IMD_Decile)) +
  geom_boxplot() +
  labs(x = "IMD Decile",
       y = "% change",
       caption = "Mean gas kWh 2010-2019, all Solent LSOas") +
  facet_wrap(la_name ~ .)

nrow(Solent)
```

#### Basic change map

```{r Mapping the change in LSOA dom mean gas comsumption}
#Green indicates a decrease in gas consumption from 2010 to 2019
#Red indicates an increase in gas consumption from 2010 to 2019

df <- dplyr::select(merged_dgas, LSOA11CD, mean_dg2010, mean_dg2019, dg_diff)
lsoa_dgas20102019 <- merge(lsoa_sf, df, by = "LSOA11CD") #merging these
ggplot2::ggplot(lsoa_dgas20102019) + geom_sf(aes(fill=dg_diff))+ 
                  scale_fill_continuous(name="Change in Gas: Mean kWh per meter", low="green",high="red")+
  labs(caption ="Solent (all MSOAs)") ##BROKEN
##Need to change the varibale names to reflect what is actually there

lsoa_dgas20102019$pc_diff <- 100*(lsoa_dgas20102019$dg_diff/lsoa_dgas20102019$mean_dg2010)
t_diff <-  dplyr::select(as.data.frame(lsoa_dgas20102019), 
                         LSOA11CD, LSOA11NM,
                         mean_dg2010, mean_dg2019, dg_diff, pc_diff)
head(arrange(t_diff, pc_diff))
head(arrange(t_diff, -pc_diff))
ggplot2::ggplot(t_diff, aes(x = pc_diff)) +
  geom_histogram()

```

#### Change map with overlay

Using leaflet

```{r Plotting leaflet for change in dom gas consumption map}

lsoa_sf_leaflet <- merge(lsoa_sf_leaflet, t_diff, by = "LSOA11CD")

qpal <- colorNumeric("Reds", lsoa_sf_leaflet$dg_diff, n=9)

leaflet(lsoa_sf_leaflet) %>%
  addTiles() %>%
  addPolygons(color = ~qpal(dg_diff),
              fillOpacity = 0.6, weight = 1.5, popup = ~(LSOA11CD),

               highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

#### Cities in more detail

```{r Southampton change in LSOA Gas}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_dgas20102019, LAD11NM == "Southampton")

# plotting the gas consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = dg_diff)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "2010-2019, Southampton LSOAs")
```
```{r Portsmouth Change in LSOA Gas}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_dgas20102019, LAD11NM == "Portsmouth")

# plotting the gas consumption of Portsmouth in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = dg_diff)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "2010-2019, Portsmouth LSOAs")
```

```{r Winchester Chnage in LSOA Gas}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_dgas20102019, LAD11NM == "Winchester")

# plotting the gas consumption of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = dg_diff)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "2010-2019, Winchester LSOAs")
```