### 2021-09-06 LSOA Deprevation.Rmd
Tweaking bits

---
title: "LSOA Deprevation maps"
author: "Meghan Kinglsey-Walsh"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # set this to FALSE if you don't want the code in the final output

# useful R libraries
library(ggplot2)
library(readr)
library(dplyr)

# GIS libraries
library(raster)
library(sf)
library(spData)
library(spDataLarge)

dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/" # default for Meghan
user <- Sys.info()[[7]] # who are we logged in as?
if(user == "ben"){
  dataFolder <- path.expand("~/University of Southampton/HCC Energy Landscape Mapping project - Documents/General/data/")
}

# this will not print anything out when knitted due to the include=FALSE setting in the chunk header
message("User: ", user)
message("dataFolder: ", dataFolder)
```

# Mapping the Lower Layer Super Output Area (LSOA)

```{r Loading in boundary data for the LSOA}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "boundaries/LSOA/lsoa_solent.shp")
message ("Loading LSOA boundaries from file")
lsoa_sf_data <- sf::read_sf(inf)
head(lsoa_sf_data)
table(lsoa_sf_data$LAD11NM)#How many LSOAs are there in each LA
```

### Mapping the Indices of Multiple Deprivation
####2015
```{r Loading in the Indices of Multiple Deprevation data 2015}
inFile <-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2015.csv")
lsoa_multidep2015 <- readr::read_csv (inFile)
lsoa_multidep2015$LSOA11CD <- lsoa_multidep2015$'lsoa11cd'
head(lsoa_multidep2015)
names(lsoa_multidep2015)
```

####2019
```{r Loading in the Indices of Multiple Deprivation data 2019}
# Multiple deprivation data at MSOA pre downloaded
inFile <-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv")
lsoa_multidep <- readr::read_csv (inFile)
lsoa_multidep$LSOA11CD <- lsoa_multidep$'lsoa11cd'
head(lsoa_multidep)
names(lsoa_multidep)
```

```{r Mapping the Indices of deprivation for the LSOA}
lsoa_multidep_merged_sf<- merge(lsoa_sf_data, lsoa_multidep)
ggplot2::ggplot(lsoa_multidep_merged_sf)+
  geom_sf(aes(fill=IMDScore))+
  scale_fill_continuous(name="IMD Score", low="green", high="red")+
  labs(caption="Solent (all LSOAs)")
```
####Map with overlay 
```{r Check leaflet for change i dom gas consumption}
library(leaflet)

st_coord_sys <- st_crs(lsoa_multidep_merged_sf)
st_coord_sys$epsg

#Transforming the coor system
if(st_coord_sys$epsg !=4326){
  map_df_trans <- st_transform(lsoa_multidep_merged_sf, "+proj=longlat +datum=WGS84")
}

```
```{r Plotting leaflet for change in dom gas consumption map}
qpal <- colorNumeric("Reds", map_df_trans$IMDScore, n=9)

leaflet(map_df_trans) %>%
  addTiles() %>%
  addPolygons(color = ~qpal(IMDScore),
              fillOpacity = 0.6, weight = 1.5, popup = ~(LSOA11CD),

               highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```
```{r Southampton LSOA Indices of dep}
#####Looking at the cities in more detail
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_multidep_merged_sf, LAD11NM == "Southampton")

# plotting the multiple scales on dep of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = IMDScore)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Southampton")
```
```{r Portsmouth LSOA Indices of dep}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_multidep_merged_sf, LAD11NM == "Portsmouth")

# plotting the change in indices of dep from 2018-2012 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = IMDScore)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Portsmouth")
```
```{r Winchester LSOA Indices of dep}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_multidep_merged_sf, LAD11NM == "Winchester")

# plotting the indices of dep 2019 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = IMDScore)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Winchester")
```
####Difference from 2015 to 2019
```{r Creating the dfference in IMD score variable}
imd_lsoa_2015 <- readr::read_csv(paste0(dataFolder, "Indices of Multiple Deprivation/File_5_ID_2015_Scores_for_the_Indices_of_Deprivation.csv")) # see https://www.gov.uk/government/statistics/english-indices-of-deprivation-2015 - this file has the actual indices in it, the other files (I don't know which you are using) have the ranks
imd_lsoa_2019 <- readr::read_csv(paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv"))
imd_lsoa_2015$imd2015 <- imd_lsoa_2015$'Index of Multiple Deprivation (IMD) Score'
imd_lsoa_2019$imd2019 <- imd_lsoa_2019$'IMDScore'

#use names(imd_lsoa_2015) to see what the column (variable) names are!
imd_lsoa_2015$LSOA11CD <-imd_lsoa_2015$`LSOA code (2011)` # <- this does!
imd_lsoa_2019$LSOA11CD <-imd_lsoa_2019$lsoa11cd

merged_imd <- merge(imd_lsoa_2015, imd_lsoa_2019, 
                    by = "LSOA11CD", # you have to use the LSOA11CD variable you just created! 'lsoa11cd' doesn't exist in these two datasets...
                    all = TRUE)
names(merged_imd)
merged_imd$imd_diff<-as.numeric(merged_imd$imd2019)-merged_imd$imd2015
summary(merged_imd)
```
```{r Mapping the change in IMD Score}
merged_imd$lsoa11cd <- merged_imd$LSOA11CD #creating a variable with the LSOA code in the same name as in sf_data - check what you created above!
lsoa_imd20152019 <- merge(lsoa_sf_data, merged_imd) #merging these
names(lsoa_imd20152019)## fixed
ggplot2::ggplot(lsoa_imd20152019) +
 geom_sf(aes(fill=imd_diff))+ 
 scale_fill_continuous(name="Change in IMD score", low="green",high="red")+
 labs(caption ="Solent (all LSOAs)")

#Histogram
lsoa_imd20152019$pc_diff <-100* (lsoa_imd20152019$imd_diff/lsoa_imd20152019$imd2015)
t_diff <-dplyr::select(as.data.frame(lsoa_imd20152019),LSOA11CD, lsoa11cd, imd2015, imd2019, imd_diff, pc_diff)
head(arrange(t_diff, pc_diff))
head(arrange(t_diff, -pc_diff))
ggplot2::ggplot(t_diff, aes(x=pc_diff)) +
  geom_histogram()
```
####Map with overlay 
```{r Check leaflet for change i dom gas consumption}
library(leaflet)

st_coord_sys <- st_crs(lsoa_imd20152019)
st_coord_sys$epsg

#Transforming the coor system
if(st_coord_sys$epsg !=4326){
  map_df_trans <- st_transform(lsoa_imd20152019, "+proj=longlat +datum=WGS84")
}

```
```{r Plotting leaflet for change in dom gas consumption map}
qpal <- colorNumeric("Reds", map_df_trans$imd_diff, n=9)

leaflet(map_df_trans) %>%
  addTiles() %>%
  addPolygons(color = ~qpal(imd_diff),
              fillOpacity = 0.6, weight = 1.5, popup = ~(LSOA11CD),

               highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

#####Looking at the cities in more detail
```{r Southampton LSOA IMD}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_imd20152019, LADnm == "Southampton") # check names of variables! Use the data file with the geometry in it
# plotting the gas consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = imd_diff)) +
  scale_fill_continuous(name = "Change in IMD score", low = "green", high = "red") +
  labs(caption = "Southampton")
#Histogram
mapData$imd_diff <-100* (mapData$imd_diff/mapData$imd2015)
t_diff <-dplyr::select(as.data.frame(lsoa_imd20152019),LSOA11CD, lsoa11cd, imd2015, imd2019, imd_diff, pc_diff)
head(arrange(t_diff, pc_diff))
head(arrange(t_diff, -pc_diff))
ggplot2::ggplot(t_diff, aes(x=pc_diff)) +
  geom_histogram()
```
```{r Portsmouth LSOA IMD}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_imd20152019, LADnm == "Portsmouth")
# plotting the gas consumption of Portsmouth in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `imd_diff`)) +
  scale_fill_continuous(name = "Change in IMD score", low = "green", high = "red") +
  labs(caption = "Portsmouth")
#Histogram
mapData$imd_diff <-100* (mapData$imd_diff/mapData$imd2015)
t_diff <-dplyr::select(as.data.frame(lsoa_imd20152019),LSOA11CD, lsoa11cd, imd2015, imd2019, imd_diff, pc_diff)
head(arrange(t_diff, pc_diff))
head(arrange(t_diff, -pc_diff))
ggplot2::ggplot(t_diff, aes(x=pc_diff)) +
  geom_histogram()
```
```{r Winchester LSOA Gas}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_imd20152019, LADnm == "Winchester")
# plotting the gas consumption of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `imd_diff`)) +
  scale_fill_continuous(name = "Change in IMD score", low = "green", high = "red") +
  labs(caption = "Winchester")
#Histogram
mapData$imd_diff <-100* (mapData$imd_diff/mapData$imd2015)
t_diff <-dplyr::select(as.data.frame(lsoa_imd20152019),LSOA11CD, lsoa11cd, imd2015, imd2019, imd_diff, pc_diff)
head(arrange(t_diff, pc_diff))
head(arrange(t_diff, -pc_diff))
ggplot2::ggplot(t_diff, aes(x=pc_diff)) +
  geom_histogram()
```
####Plotting the relationship between deprevation and Fuel Poverty