### 2021-09-14 mappingParishes.Rmd
added annual change plots

---
title: "Mapping parishe emissions: Framlingham as an example"
author: "Ben Anderson"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # set this to FALSE if you don't want the code in the final output

# Libraries ----
# useful R libraries
library(ggplot2)
library(dplyr)
library(data.table)

# GIS libraries
library(leaflet)
library(raster)
library(sf)

# Paramaters ----

dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/" # default for Meghan
user <- Sys.info()[[7]] # who are we logged in as?
if(user == "ben"){
  dataFolder <- path.expand("~/Dropbox/data/")
}

# this will not print anything out when knitted due to the include=FALSE setting in the chunk header
message("User: ", user)
message("dataFolder: ", dataFolder)

# Functions ----

selectSoton <- function(dt){
  # function we can re-use
  # assumes a data.table
  # assumes LA name = la_name (may need to make new var first)
  # returns just the rows that have Local Authority == Southampton
  select_dt <- dt[la_name %like% "Southampton"]
  return(select_dt)
}

selectSolent <- function(dt){
  # function we can re-use
  # assumes a data.table
  # assumes LA name = la_name (may need to make new var first)
  # returns just the rows that have Local Authority in Solent area
  select_dt <- dt[la_name == "Southampton" | 
                   la_name == "Portsmouth" |
                   la_name == "Winchester" |
                   la_name == "Eastleigh" |
                   la_name == "Isle of Wight" |
                   la_name == "Fareham" |
                   la_name == "Gosport" |
                   la_name == "Test Valley" |
                   la_name == "East Hampshire" |
                   la_name == "Havant" |
                   la_name == "New Forest" |
                   la_name == "Hart" |
                   la_name == "Basingstoke and Deane"]
  return(select_dt)
}

selectParish <- function(dt, parish = parish){
   # as above but selects a parish based on name
  select_dt <- dt[parncp19nm == get(parish)]
}

# Data ----
f <- "lookups/Lower_Layer_Super_Output_Area_(2011)_to_Ward_(2020)_to_LAD_(2020)_Lookup_in_England_and_Wales_V2.csv.gz"
#lsoa_lut <- data.table::fread(paste0(dataFolder, f))
lsoa_lookup <- data.table::fread(paste0(dataFolder, "lookups/lsoa_lookup.csv.gz")) # reduced version

parish_lookup <- data.table::fread(paste0(dataFolder, "lookups/Parish_to_Ward_to_Local_Authority_District__December_2020__Lookup_in_England_and_Wales.csv.gz"))

# 
```

# Intro

Focus on Framlingham (for now, intended to be gneric)

# Mapping parishes

Example: Framlingham

```{r loadParishShapefile}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "EW_parishes/Parishes_and_Non_Civil_Parished_Areas_(April_2019)_EW_BGC/Parishes_and_Non_Civil_Parished_Areas_(April_2019)_EW_BGC.shp")
message ("Loading parish boundaries from file")
parish_sf_data <- sf::read_sf(inf)
names(parish_sf_data)
```

Build a simple map just to check the size of the parish (Figure \@ref(fig:simpleParishMap).

```{r simpleParishMap, fig.cap = "LSOA check map (shows MSOA and ward names when clicked"}

st_coord_sys <- sf::st_crs(parish_sf_data) # check coord system
#st_coord_sys # current coord system EPSG: 4326 (is what leaflet wants - good)

# transform the coord system if required
if(st_coord_sys$epsg != 4326){
 parish_sf_trans <- st_transform(parish_sf_data, "+proj=longlat +datum=WGS84")
}

fram <- dplyr::filter(parish_sf_trans, parncp19nm == "Framlingham")

leaflet(fram) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(fillOpacity = 0.2, weight = 1.5, popup = ~(paste0("Name ", parncp19nm,
                      "<br> LA: " ,lad19nm)
                      ), # popups clicked
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

Now build a simple map from LSOAs based on the Framlingham electoral ward to see how well it matches the parish (Figure \@ref(fig:simpleLSOAMap).

```{r loadLSOAShapefile}
inf<-paste0(dataFolder, "/UK_census2011/Lower_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3-shp/Lower_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3.shp")
message ("Loading LSOA boundaries from file")
lsoa_sf_data <- sf::read_sf(inf)
names(lsoa_sf_data)

# merge the LSOA look-up table
df <- merge(lsoa_sf_data, lsoa_lookup, by = "LSOA11CD")

fram_lsoa_sf_data <- dplyr::filter(df, WD20NM %like% "Framlingham")

names(fram_lsoa_sf_data)
```

```{r simpleLSOAMap, fig.cap = "LSOA check map (shows MSOA and ward names when clicked"}

st_coord_sys <- sf::st_crs(fram_lsoa_sf_data) # check coord system
#st_coord_sys # current coord system EPSG: 4326 (is what leaflet wants - good)

# transform the coord system if required
if(st_coord_sys$epsg != 4326){
 fram_lsoa_sf_trans <- st_transform(fram_lsoa_sf_data, "+proj=longlat +datum=WGS84")
}

leaflet(fram_lsoa_sf_trans) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(fillOpacity = 0.2, weight = 1.5, popup = ~(paste0("LSOA Name ", LSOA11NM.x,
                      "<br> LSOA code:: " ,LSOA11CD,
                      "<br> LA: " ,LA11NM)
                      ), # popups clicked
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

dplyr::select(fram_lsoa_sf_data, "LSOA11CD","LSOA11NM.x", "MSOA11CD","MSOA11NM", "WD20NM", )
```

This map suggests that two LSOAs:

 * E01030172 Suffolk Coastal 002B
 * E01030173 Suffolk Coastal 002C

Are a close match to the Framlingham civil parish shown in the first map. This will enable us to extract residential gas & electricity use.

Commercial gas & electricity use is at MSOA level. This may exceed the parish boundaries - find E02006288 (see above) and map it - Figure \@ref(fig:simpleMSOAMap).

```{r loadMSOAShapefile}

inf<-paste0(dataFolder, "/UK_census2011/Middle_Layer_Super_Output_Areas__December_2011__Boundaries-shp/Middle_Layer_Super_Output_Areas__December_2011__Boundaries.shp")
message ("Loading MSOA boundaries from file")
msoa_sf_data <- sf::read_sf(inf)
names(msoa_sf_data)

fram_msoa_sf_data <- dplyr::filter(msoa_sf_data, msoa11cd == "E02006288")

names(fram_msoa_sf_data)
```

```{r simpleMSOAMap, fig.cap = "MSOA check map (shows MSOA and ward names when clicked"}

st_coord_sys <- sf::st_crs(fram_msoa_sf_data) # check coord system
#st_coord_sys # current coord system EPSG: 4326 (is what leaflet wants - good)

# transform the coord system if required
if(st_coord_sys$epsg != 4326){
 fram_msoa_sf_trans <- st_transform(fram_msoa_sf_data, "+proj=longlat +datum=WGS84")
}

leaflet(fram_msoa_sf_trans) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(fillOpacity = 0.2, weight = 1.5, popup = ~(paste0("MSOA Name ", msoa11nm)), # popups clicked
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

Yep, way too big. In fact this looks much like the [Framlingham electoral ward](https://www.suffolk.gov.uk/assets/council-and-democracy/how-the-council-works/SCC-Electoral-Division-Ward-Parish-map.pdf)...

This means we can only get residential electricity & gas at LSOA level and aggregate, all other data is at MSOA or LA level. See CREDS [carbon.place](https://www.carbon.place/) for LSOA emissions estimates or CSE's impact tool for parish/LA emissions - both of these use the LSOA data for 2018/19 as part of their inputs.

# Energy use

2010 & 2019 residential gas & electricity consumption data from:

 * https://www.gov.uk/government/statistics/lower-and-middle-super-output-areas-gas-consumption
 * https://www.gov.uk/government/statistics/lower-and-middle-super-output-areas-electricity-consumption

Define Framlingham by selecting the following LSOAs:

 * E01030172 Suffolk Coastal 002B
 * E01030173 Suffolk Coastal 002C

## Gas 2010

Note the 'old' local authority label.

```{r getResidentialGas2010}
gas2010_dt <- data.table::fread(paste0(dataFolder, "beis/subnationalGas/lsoaDom/LSOA_GAS_2010.csv.gz"))
gas2010_dt[, LSOA11CD := `Lower Layer Super Output Area (LSOA) Code`]
gas2010_dt[, total_gas_kWh := `Consumption (kWh)`]
setkey(gas2010_dt, LSOA11CD)
setkey(lsoa_lookup, LSOA11CD)
gas2010_dt <- gas2010_dt[lsoa_lookup]
fram_gas2010_dt <- gas2010_dt[LSOA11CD == "E01030172" |
                        LSOA11CD == "E01030173"]

fram_gas2010_dt[, .(`LSOA Name`, LSOA11CD, WD20NM, `Number of meters`, total_gas_kWh)]

message("Framlingham: Total number of gas meters = ", sum(fram_gas2010_dt$`Number of meters`))

message("Total residential gas kWh = ", sum(fram_gas2010_dt$total_gas_kWh))
```

Using BEIS gas emissions factors for 2010 (https://www.gov.uk/government/publications/greenhouse-gas-reporting-conversion-factors-2010) doesn't help - no gas value. Use 2015 -   0.18445 kg CO2e/kWh (it's roughly constant over time in any case)

```{r residentialGasCO2e2010}
fram_gas2010_dt[, T_co2e_gas := (total_gas_kWh * 0.18445)/1000]

fram_gas2010_dt[, .(`LSOA Name`, LSOA11CD, WD20NM, `Number of meters`, total_gas_kWh, T_co2e_gas)]

message("Total T CO2e from gas  = ", sum(fram_gas2010_dt$T_co2e_gas))
```

## Electricity 2010

  BEIS CI: 0.48531 kg CO2e/kWh - very high due to use of goal.

```{r getResidentialElec2010}
elec2010_dt <- data.table::fread(paste0(dataFolder, "beis/subnationalElec/lsoaDom/LSOA_ELEC_2010.csv.gz"))
elec2010_dt[, LSOA11CD := `Lower Layer Super Output Area (LSOA) Code`]
elec2010_dt[, total_elec_kWh := `Total domestic electricity consumption (kWh)`]
setkey(elec2010_dt, LSOA11CD)
setkey(lsoa_lookup, LSOA11CD)
elec2010_dt <- elec2010_dt[lsoa_lookup]
fram_elec2010_dt <- elec2010_dt[LSOA11CD == "E01030172" |
                        LSOA11CD == "E01030173"]

fram_elec2010_dt[, .(LSOA11NM, LSOA11CD, WD20NM, `Total number of domestic electricity meters`, total_elec_kWh)]

message("Total number of elec meters = ", sum(fram_elec2010_dt$`Total number of domestic electricity meters`))

message("Total residential elec kWh = ", sum(fram_elec2010_dt$total_elec_kWh))
```
The difference between number of gas and electricity meters more or less shows those off-gas - but there could be two electricity meters per house - e.g. anyone using Economy-7.

```{r residentialElecCO2e2010}
fram_elec2010_dt[, T_co2e_elec := (total_elec_kWh * 0.48531)/1000]

fram_elec2010_dt[, .(LSOA11NM, LSOA11CD, WD20NM, `Total number of domestic electricity meters`, total_elec_kWh, T_co2e_elec)]

s2010 <- sum(fram_elec2010_dt$T_co2e_elec)
s2021 <- (sum(fram_elec2010_dt$total_elec_kWh) *   0.21233 )/1000

message("Total T CO2e from electricity  = ", s2010)

message("Total T CO2e from this much electricity in 2021 using the most recent grid CI factor would be = ", s2021)

message("That's ", 100*round(s2021/s2010,2), " % lower.")
```

## Gas 2019 (latest data, pre-covid)

```{r getResidentialGas2019}
gas_dt <- data.table::fread(paste0(dataFolder, "beis/subnationalGas/lsoaDom/LSOA_GAS_2019.csv.gz"))
gas_dt[, LSOA11CD := `Lower Layer Super Output Area (LSOA) Code`]
gas_dt[, total_gas_kWh := `Consumption (kWh)`]
setkey(gas_dt, LSOA11CD)
setkey(lsoa_lookup, LSOA11CD)
gas_dt <- gas_dt[lsoa_lookup]
fram_gas2019_dt <- gas_dt[LSOA11CD == "E01030172" |
                        LSOA11CD == "E01030173"]

fram_gas2019_dt[, .(`LSOA Name`, LSOA11CD, WD20NM, `Number of consuming meters`, total_gas_kWh)]

message("Total number of consuming gas meters 2019 = ", sum(fram_gas2019_dt$`Number of consuming meters`))

message("Total residential gas kWh 2019 = ", sum(fram_gas2019_dt$total_gas_kWh))
```

Using BEIS gas emissions factors for 2019 (https://www.gov.uk/government/publications/greenhouse-gas-reporting-conversion-factors-2019).

  0.18385 kg CO2e/kWh +   0.02391 (WTT)

```{r residentialGasCO2e2019}
fram_gas2019_dt[, T_co2e_gas := (total_gas_kWh * (0.18385 + 0.02391))/1000]

fram_gas2019_dt[, .(`LSOA Name`, LSOA11CD, WD20NM, `Number of consuming meters`, total_gas_kWh, T_co2e_gas)]

message("Total T CO2e from gas  = ", sum(fram_gas2019_dt$T_co2e_gas))

```

## Electricity 2019 (latest data, pre-covid)

  BEIS CI:   0.2556   kg CO2e/kWh +   0.03565  + 0.02170 (WTT + T&D)

```{r getResidentialElec2019}
elec2019_dt <- data.table::fread(paste0(dataFolder, "beis/subnationalElec/lsoaDom/LSOA_ELEC_2019.csv.gz"))
elec2019_dt[, LSOA11CD := `Lower Layer Super Output Area (LSOA) Code`]
elec2019_dt[, total_elec_kWh := `Total domestic electricity consumption (kWh)`]
setkey(elec2019_dt, LSOA11CD)
setkey(lsoa_lookup, LSOA11CD)
elec2019_dt <- elec2019_dt[lsoa_lookup]
fram_elec2019_dt <- elec2019_dt[LSOA11CD == "E01030172" |
                        LSOA11CD == "E01030173"]

fram_elec2019_dt[, .(LSOA11NM, LSOA11CD, WD20NM, `Total number of domestic electricity meters`, total_elec_kWh)]

message("Total number of elec meters = ", sum(fram_elec2019_dt$`Total number of domestic electricity meters`))

message("Total residential elec kWh = ", sum(fram_elec2019_dt$total_elec_kWh))
```
The difference between number of gas and electricity meters more or less shows those off-gas - but there could be two electricity meters per house - e.g. anyone using Economy-7.

```{r residentialElecCO2e2019}
fram_elec2019_dt[, T_co2e_elec := (total_elec_kWh * (0.2556 + 0.03565 + 0.02170))/1000]

fram_elec2019_dt[, .(LSOA11NM, LSOA11CD, WD20NM, `Total number of domestic electricity meters`, total_elec_kWh, T_co2e_elec)]

message("Total T CO2e from electricity  = ", sum(fram_elec2019_dt$T_co2e_elec))

```

## 2010 - 2019 change

### Gas 2010-2019

```{r gasChange}

# add to 2010
fram_gas2010_dt[, year := 2010]
fram_gas2019_dt[, year := 2019]

fg_2010 <- fram_gas2010_dt[,.(LSOA11CD, LSOA11NM, WD20NM, total_gas_kWh, T_co2e_gas, nMeters = `Number of meters`, year)]
fg_2019 <- fram_gas2019_dt[ ,.(LSOA11CD, LSOA11NM, WD20NM, total_gas_kWh, T_co2e_gas, nMeters = `Number of consuming meters`, year)]
fram_gas_dt <- rbind(fg_2010, fg_2019)

ggplot2::ggplot(fram_gas_dt, aes(x = year, y = total_gas_kWh/1000, colour = LSOA11NM)) +
  geom_line() +
  geom_point() +
  labs(y = "Gas MWh")

ggplot2::ggplot(fram_gas_dt, aes(x = year, y = T_co2e_gas, colour = LSOA11NM)) +
  geom_line() +
  geom_point()

fram_gas_dt

gas_meters_diff = sum(fram_gas2019_dt$`Number of consuming meters`)-sum(fram_gas2010_dt$`Number of meters`)
gas_meters_diff_pc <- 100*(gas_meters_diff/sum(fram_gas2010_dt$`Number of meters`))
message("Change in number of meters = ", gas_meters_diff , " (", round(gas_meters_diff_pc)," %)")

gas_diff = sum(fram_gas2019_dt$total_gas_kWh)-sum(fram_gas2010_dt$total_gas_kWh)
pc_gas_diff <- 100*(gas_diff/sum(fram_gas2019_dt$total_gas_kWh))
message("Change in total gas use = ", gas_diff/1000 , " MWh")
message("% change in total gas use = ", round(pc_gas_diff,2), " %")

gas_Tco2e_diff = sum(fram_gas2019_dt$T_co2e_gas)-sum(fram_gas2010_dt$T_co2e_gas)
pc_gas_Tco2e_diff <- 100 * (gas_Tco2e_diff/sum(fram_gas2010_dt$T_co2e_gas))
message("% change in T CO2e from gas use = ", round(pc_gas_Tco2e_diff,2), " %")
```

```{r getAllGasFiles}
# slightly more complex - need to load all files & select LSOAs
getGasFiles <- function(files){
  dt <- data.table::data.table()
  for(f in files){
    d <- data.table::fread(f)
    message("Loading ", f)
    names(d)
    s <- strsplit(f, "_")
    suffix <- strsplit(s[[1]][3], ".", fixed = TRUE)
    year <- suffix[[1]][1]
    d[, year := year]
    # if these variable names vary by year have to set here
    d[, total_gas_kWh := `Consumption (kWh)`]

    dt <- rbind(dt,d, fill=TRUE) # prevents breakage when new var name introduced in 2016
  }

  return(dt)
}

files <- list.files(paste0(dataFolder, "beis/subnationalGas/lsoaDom"), 
                    pattern = ".gz", full.names = TRUE)

gas_allYears <- getGasFiles(files)

gas_allYears[, total_nMeters := ifelse(is.na(`Number of meters`), # must be after 2014
                                       `Number of consuming meters`, # so this is true
                                       `Number of meters`)] # pre 2015

gas_allYears[, .(nMeters = mean(`Number of meters`),
             nConsumingMeters = mean(`Number of consuming meters`),
             nAllMeters = mean(total_nMeters)), 
             keyby = .(year)]

names(gas_allYears)
gas_allYears[, LSOA11CD := `Lower Layer Super Output Area (LSOA) Code`]

```

```{r plotGasOverTime}

fram_gas_allYears <- gas_allYears[LSOA11CD == "E01030172" |
                        LSOA11CD == "E01030173"]

ggplot2::ggplot(fram_gas_allYears, aes(x = year, y = total_nMeters, 
                                        group = LSOA11CD,
                                        colour = LSOA11CD)) +
  geom_line() +
  geom_point() +
  labs(y = "Total number of gas meters")

ggplot2::ggplot(fram_gas_allYears, aes(x = year, y = total_gas_kWh/1000, 
                                        group = LSOA11CD,
                                        colour = LSOA11CD)) +
  geom_line() +
  geom_point() +
  labs(y = "Total residential gas (MWh)")

ggplot2::ggplot(fram_gas_allYears, aes(x = year, y = total_gas_kWh/total_nMeters, 
                                        group = LSOA11CD,
                                        colour = LSOA11CD)) +
  geom_line() +
  geom_point() +
  labs(y = "Mean residential gas kWh per meter",
       caption = "This is an indicator of increasing energy efficiency over time")
```

### Electricity 2010-2019

```{r elecChange}

# add to 2010
fram_elec2010_dt[, year := 2010]
fram_elec2019_dt[, year := 2019]

fe_2010 <- fram_elec2010_dt[,.(LSOA11CD, LSOA11NM, WD20NM, total_elec_kWh, T_co2e_elec, nMeters = `Total number of domestic electricity meters`, year)]
fe_2019 <- fram_elec2019_dt[ ,.(LSOA11CD, LSOA11NM, WD20NM, total_elec_kWh, T_co2e_elec, nMeters = `Total number of domestic electricity meters`, year)]
fram_elec_dt <- rbind(fe_2010, fe_2019)

ggplot2::ggplot(fram_elec_dt, aes(x = year, y = total_elec_kWh/1000, colour = LSOA11NM)) +
  geom_line() +
  geom_point() +
  labs(y = "Electricity MWh")

ggplot2::ggplot(fram_elec_dt, aes(x = year, y = T_co2e_elec, colour = LSOA11NM)) +
  geom_line() +
  geom_point()

fram_elec_dt

elec_meters_diff = sum(fram_elec2019_dt$`Total number of domestic electricity meters`)-sum(fram_elec2010_dt$`Total number of domestic electricity meters`)
elec_meters_diff_pc <- 100*(elec_meters_diff/sum(fram_elec2010_dt$`Total number of domestic electricity meters`))
message("Change in number of meters = ", elec_meters_diff , " (", round(elec_meters_diff_pc)," %) - should probably be about the same as number of gas meters until new builds with electricity only are more common")

elec_diff = sum(fram_elec2019_dt$total_elec_kWh)-sum(fram_elec2010_dt$total_elec_kWh)
pc_elec_diff <- 100*(elec_diff/sum(fram_elec2010_dt$total_elec_kWh))
message("Change in total electricity use = ", elec_diff/1000 , " MWh")
message("% change in total electricity use = ", round(pc_elec_diff,2), " %")

elec_Tco2e_diff = sum(fram_elec2019_dt$T_co2e_elec)-sum(fram_elec2010_dt$T_co2e_elec)
pc_elec_Tco2e_diff <- 100 * (elec_Tco2e_diff/sum(fram_elec2010_dt$T_co2e_elec))
message("% change in T CO2e from electricity use = ", round(pc_elec_Tco2e_diff,2), " %")
```

```{r getAllElecFiles}
# slightly more complex - need to load all files & select LSOAs
getElecFiles <- function(files){
  dt <- data.table::data.table()
  for(f in files){
    d <- data.table::fread(f)
    message("Loading ", f)
    names(d)
    s <- strsplit(f, "_")
    suffix <- strsplit(s[[1]][3], ".", fixed = TRUE)
    year <- suffix[[1]][1]
    d[, year := year]
    # if these variable names vary by year have to set here
    d[, total_elec_kWh := `Total domestic electricity consumption (kWh)`]
    d[, total_nMeters := `Total number of domestic electricity meters`]
    dt <- rbind(dt,d)
  }

  return(dt)
}

files <- list.files(paste0(dataFolder, "beis/subnationalElec/lsoaDom"), 
                    pattern = ".gz", full.names = TRUE)

elec_allYears <- getElecFiles(files)
names(elec_allYears)
elec_allYears[, LSOA11CD := `Lower Layer Super Output Area (LSOA) Code`]
elec_allYears[, LSOA11NM := `Lower Layer Super Output Area (LSOA) Name`]

```

```{r plotElecOverTime}

fram_elec_allYears <- elec_allYears[LSOA11CD == "E01030172" |
                        LSOA11CD == "E01030173"]

ggplot2::ggplot(fram_elec_allYears, aes(x = year, y = total_nMeters, 
                                        group = LSOA11CD,
                                        colour = LSOA11CD)) +
  geom_line() +
  geom_point() +
  labs(y = "Total number of electricity meters")

ggplot2::ggplot(fram_elec_allYears, aes(x = year, y = total_elec_kWh/1000, 
                                        group = LSOA11CD,
                                        colour = LSOA11CD)) +
  geom_line() +
  geom_point() +
  labs(y = "Total residential electricity (MWh)")

ggplot2::ggplot(fram_elec_allYears, aes(x = year, y = total_elec_kWh/total_nMeters, 
                                        group = LSOA11CD,
                                        colour = LSOA11CD)) +
  geom_line() +
  geom_point() +
  labs(y = "Mean residential kWh per meter",
       caption = "This is an indicator of increasing energy efficiency over time")
```

# The end

```{r environment}
R.Version()

```